### 2017-11-16 paper/manuscript.tex
Create manuscript template

% paper for Zone based traffic assignment algorithm. Short 4 page paper for ICT Express issue
% The aim here is to show an increase in scalibility with a preservation of traffic reduction (also note that the zone system will work with any model)

% for condensed version of paper, all pseudocode is dropped

% TODO:
% launch smaller zone experiments (should not take too long) (zone size of 5 is extremely slow) 
% generate graphs
% final passthrough
% done

\documentclass[conference]{IEEEtran}
\usepackage{mathtools}

\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx

\usepackage{graphicx}
\graphicspath{ {images/} }

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\raggedbottom

\title{An Improved Sequential Data Mining Model for Hurricane Trajectory Prediction}
% A novel algorithm for scalable traffic congestion reduction
% An adaptible model for scalability in traffic assignment algorithms
% Z-BAR: A Zone-Based traffic Assignment Algorithm for scalable congestion Reduction

% Is this an Algorithm or a framework: We want to highlight compatability

\author{\IEEEauthorblockN{Jasmin Bissonnette, Caden Marofke, Taylor Cox}
\IEEEauthorblockA{Department of Computer Science\\
University of Manitoba\\
Winnipeg, Manitoba\\
Email: \{bissonnj, marofkec, coxt3\}@myumanitoca.}}

\maketitle

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
Hurricane Trajectory, Apriori, Prediction
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\section{Introduction}

\section{Related Work}

% Methodology and Solution go here

\section{Experimental Setup}

% Need CSV output for metrics

\section{Experimental Results}

\section{Conclusions and Future Work} % what we found:

\begin{thebibliography}{1}
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\bibitem{} 
\end{thebibliography}

\end{document}